---
title: "Homework 2"
author: "Deepika Dilip"
date: "3/20/2019"
output:
  pdf_document: default
  html_document: default
---
```{r}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning=F)
knitr::opts_chunk$set(message=F)
```

```{r}
library(caret) # only for plot 
library(splines) 
library(lasso2) # only for data 
library(mgcv) 
library(tidyverse) 
library(ggplot2)
library(janitor)
library(boot) #for CV
```

#Importing the Data

```{r}
setwd('~/Dropbox/Spring 2019/Data Science II/Homework')
concrete <-read.csv("concrete.csv")
concrete= janitor::clean_names(concrete)
```

#Defining the predictor and outcome matrices
```{r}

# matrix of predictors 
x <- model.matrix(compressive_strength~.,concrete)[,-1] 
# vector of response 
y <- concrete$compressive_strength
```

#Part A: Plotting the Data
```{r}
theme1 <- trellis.par.get() 
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5) 
theme1$plot.symbol$pch <- 16 
theme1$plot.line$col <- rgb(.8, .1, .1, 1) 
theme1$plot.line$lwd <- 2 
theme1$strip.background$col <- rgb(.0, .2, .6, .2) 
trellis.par.set(theme1) 
featurePlot(x, y, plot = "scatter", labels = c("","Compressive Strength"), type = c("p"), layout = c(4, 2))
```

#Part B: Polynomial regression
```{r}
fit1 <- lm(compressive_strength~water, data = concrete) 
fit2 <- lm(compressive_strength~poly(water,2), data = concrete) 
fit3 <- lm(compressive_strength~poly(water,3), data = concrete) 
fit4 <- lm(compressive_strength~poly(water,4), data = concrete) 

cv.error=rep (0,5)
for (i in 1:5){
  glm.fit=glm(compressive_strength~poly(water,i), data=concrete) 
  cv.error[i] = cv.glm(concrete, glm.fit)$delta[1]
}
print(cv.error)
```

```{r}
anova(fit1,fit2,fit3,fit4)
```

Based on cross validation, we want the d that minimizes the cross validation error. Therefore we should use d=4.

```{r}
anova(fit1,fit2,fit3,fit4)
```

```{r}
pred.fit1 <- predict(fit1, x = water)
pred.fit2 <- predict(fit2, x = water)
pred.fit3 <- predict(fit3, x = water)
pred.fit4 <- predict(fit4, x = water)

p1 <- ggplot(data=concrete) + 
  geom_point(aes(x=water, y=compressive_strength), color="black", size=0.5) +
  geom_line(aes(x = water, y = pred.fit1),
            color = rgb(.3, .1, .1, 1)) +
  geom_line(aes(x = water, y = pred.fit2),
            color = 'blue') +
  geom_line(aes(x = water, y = pred.fit3),
            color = 'green') +
  geom_line(aes(x = water, y = pred.fit4),
            color = 'red')
print(p1)
```

#Part C: Smoothing spline
```{r}
fit.ss <- smooth.spline(concrete$water, concrete$compressive_strength) 
fit.ss$df

waterlimits <- range(concrete$water) 
water.grid <- seq(from = waterlimits[1],to = waterlimits[2])

pred.ss <- predict(fit.ss, x = water.grid)
pred.ss.df <- data.frame(pred = pred.ss$y, water = water.grid)

spline_plot <- ggplot(data = concrete, aes(x = water, y = compressive_strength)) + 
  geom_point(color = rgb(.2, .4, .2, .5)) + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df, color = rgb(.8, .1, .1, 1)) + theme_bw()

print(spline_plot)
```
##69 degrees of freedom


##Range of df: 2-12
```{r}
par(mfrow=c(3,4))
for (i in 2:12) {
  fit.ss1 = smooth.spline(concrete$water, concrete$compressive_strength, df = i)
  pred.ss_df <- predict(fit.ss1, x = water.grid)
  pred.ss_df <- data.frame(pred = pred.ss_df$y, water = water.grid)
plot(concrete$water, concrete$compressive_strength, cex = .5, col = "grey") + title(paste("Degrees of freedom = ", round(fit.ss1$df)), outer = F) + lines(water.grid, pred.ss_df$pred, lwd = 2, col = "green")
  }

```

Based on the results, it seems that df = 69 is most effective.

#Part D: GAM
```{r}
gam.m1 <- gam(compressive_strength ~ cement + blast_furnace_slag
              + fly_ash + water + superplasticizer + coarse_aggregate 
              + fine_aggregate + age, data = concrete)
gam.m2 <- gam(compressive_strength ~ cement + blast_furnace_slag
              + fly_ash + s(water) + s(superplasticizer) + coarse_aggregate 
              + fine_aggregate + s(age), data = concrete)
anova(gam.m1, gam.m2, test = "F")

#Plotting the Model
par(mfrow=c(1,3)) #to partition the Plotting Window
plot(gam.m2,se = TRUE) 
```

Based on the scatter plots from Question 1, we can assume age, water, and superplasticizer require smoothing as they are nonlinear.